---
title: "HUANG"
format: html
editor: visual
---
```{r}
here::i_am("grades-test.Rproj")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  message = FALSE,
  warning = FALSE
)
```

```{r}
library(here)
library(vroom)
library(dplyr)
library(ggplot2)
library(tidyr)
library(knitr)
```

# Q1 Load data
```{r}
courses <- vroom(here("data","courses.csv"))
```


# Q2 Formatted table

```{r}
courses |>
  arrange(courses) |>
  kable(
    caption = "List of courses",
    align   = "l"
  )
```


# Q3 Load Students dataset

```{r}

students <- vroom(
  here("data", "students.csv"),
  col_types = cols(
    id         = col_integer(),
    group      = col_factor(),
    birth_date = col_date(),
    sex        = col_factor()
  )
)
```

# Q4 Load grades dataset 

```{r}
grades <- vroom(
  here("data", "grades.csv"),
  delim = ":",
  na    = "*"
)
```

# Q5 Number of students per group

```{r}
students |>
  count(group, name = "n_students") |>
  ggplot(aes(x = group, y = n_students)) +
  geom_col() +
  labs(
    title = "Number of students per group",
    x = "Group",
    y = "Number of students"
  ) +
  theme_minimal()
```

# Q6 Gender balance in each group

```{r}
students |>
  ggplot(aes(x = group, fill = sex)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(
    title = "Gender balance by group",
    x = "Group",
    y = "Share of students",
    fill = "Sex"
  ) +
  theme_minimal()
```

# Q7 Distribution of the age of the student
```{r}
students_age <- students |>
  mutate(
    age = as.numeric(difftime(Sys.Date(), birth_date, units = "days")) / 365.25
  )

students_age |>
  ggplot(aes(x = age)) +
  geom_histogram(binwidth = 1) +
  labs(
    title = "Distribution of students' age",
    x = "Age (years)",
    y = "Number of students"
  ) +
  theme_minimal()
```


# Q8 Median age of each student in each group
```{r}
median_age_by_group <- students_age |>
  group_by(group) |>
  summarise(
    median_age = as.integer(round(median(age, na.rm = TRUE))),
    .groups = "drop"
  )

median_age_by_group |>
  kable(
    caption = "Median age of students by group",
    col.names = c("Group", "Median age")
  )
```

# Q9 Oldest student in each group
```{r}
oldest_by_group <- students_age |>
  group_by(group) |>
  slice_max(age, n = 1, with_ties = FALSE) |>
  ungroup() |>
  arrange(desc(age)) |>
  transmute(
    id,
    sex,
    age = round(age, 1) 
  )


oldest_by_group |>
  kable(
    caption   = "Oldest student in each group",
    col.names = c("ID", "Sex", "Age")
  )
```

# Q10 Number of grades in the dataset
```{r}
nb_grades <- grades |>
  filter(!is.na(grade)) |>
  nrow()
```

The data set contains `r nb_grades` grades.


# Q11 Average of all the grades of one course

```{r}
neuro_grades_by_group <- grades |>
  inner_join(courses, by = "course_id") |>
  filter(course == "Neurotechnology and Mind Control") |>
  inner_join(students |> select(id, group), by = "id") |>
  group_by(group) |>
  summarise(
    mean_grade = mean(grade, na.rm = TRUE),
    .groups = "drop"
  )

neuro_grades_by_group |>
  ggplot(aes(x = group, y = mean_grade)) +
  geom_col() +
  labs(
    title = "Average grade in Neurotechnology and Mind Control by group",
    x = "Group",
    y = "Average grade"
  )
```


# Q12 Distribution of the grades of 2 semesters
```{r}
grades_sem <- grades |>
  inner_join(courses |> select(course_id, semester), by = "course_id") |>
  filter(!is.na(grade)) |>
  mutate(semester = as.factor(semester))

grades_sem |>
  ggplot(aes(x = grade, fill = semester)) +
  geom_density(alpha = 0.5) +
  labs(
    title = "Distribution of grades by semester",
    x = "Grade",
    y = "Density",
    fill = "Semester"
  )
```

# Q13 Number of grades per students
```{r}
grades_per_student <- grades |>
  filter(!is.na(grade)) |>
  count(id, name = "n_grades")

students_grades <- students |>
  left_join(grades_per_student, by = "id") |>
  mutate(n_grades = replace_na(n_grades, 0L))

students_grades |>
  head(10) |>
  kable(
    caption = "Extract of the number of grades per student",
    col.names = c("ID", "Group", "Birth date", "Sex", "Number of grades")
  )

summary_grades <- students_grades |>
  summarise(
    min    = min(n_grades),
    max    = max(n_grades),
    mean   = mean(n_grades),
    median = median(n_grades)
  )

summary_grades |>
  kable(
    caption = "Summary of the number of grades per student",
    digits = 2,
    col.names = c("Minimum", "Maximum", "Average", "Median")
  )
```

# Q14 Distribution of the grades of male and female students
```{r}
students_grades |>
  ggplot(aes(x = n_grades)) +
  geom_histogram(binwidth = 1) +
  facet_wrap(~ sex) +
  labs(
    title = "Distribution of the number of grades by sex",
    x = "Number of grades",
    y = "Number of students"
  ) +
  theme_minimal()
```

# Q15 Dataframe course Astroengineering and Planetary Colonization 
```{r}
astro_grades <- grades |>
  inner_join(courses, by = "course_id") |>
  filter(course == "Astroengineering and Planetary Colonization") |>
  filter(!is.na(grade)) |>
  count(id, name = "n_grades_astro")

astro_grades_per_student <- students |>
  select(id, group) |>
  left_join(astro_grades, by = "id") |>
  mutate(n_grades_astro = replace_na(n_grades_astro, 0L))

astro_grades_per_student |>
  head(10) |>
  kable(
    caption = "Number of grades in Astroengineering and Planetary Colonization per student",
    col.names = c("ID", "Group", "Number of grades in Astroengineering")
  )
```

# Q16 Distribution of the number of grades in Astroengineering
```{r}
dist_astro_grades <- astro_grades_per_student |>
  count(n_grades_astro, name = "n_students")

dist_astro_grades |>
  kable(
    caption   = "Distribution of the number of grades in Astroengineering",
    col.names = c("Number of grades", "Number of students")
  )

dist_astro_grades |>
  ggplot(aes(x = n_grades_astro, y = n_students)) +
  geom_col() +
  labs(
    title = "Distribution of the number of grades in Astroengineering",
    x = "Number of grades",
    y = "Number of students"
  ) +
  theme_minimal()
```

# Q17 Number of grades depending on the group
```{r}
astro_grades_per_student |>
  mutate(group = as.factor(group)) |>
  ggplot(aes(x = group, y = n_grades_astro)) +
  geom_violin(trim = FALSE) +
  labs(
    title = "Distribution of Astroengineering grades per student by group",
    x = "Group",
    y = "Number of Astroengineering grades"
  ) +
  theme_minimal()
```


# Q18 Dependency of the number of grades with respect to both the group and the sex of the students
```{r}
astro_group_sex <- astro_grades_per_student |>
  left_join(students |>
              select(id, sex),
            by = "id") |>
  mutate(
    group = as.factor(group),
    sex   = as.factor(sex)
  )

astro_group_sex |>
  ggplot(aes(x = group, y = n_grades_astro, fill = sex)) +
  geom_violin(trim=FALSE) +
  labs(
    title = "Number of Astroengineering grades per student by group and sex",
    x = "Group",
    y = "Number of Astroengineering grades",
    fill = "Sex"
  ) +
  theme_minimal()
```

